```{r load-data}

library(tidyverse)
library(magrittr)                       # fancy pipes like %T>%

# loads "d", the data_frame we generated last time
d <- readr::read_rds('../data/batch2_33final-clean.rds')
```

# Specifying a statistical model

In R, statistical models are commonly specified using a _formula_, which
describes the relationship between _dependent_ (observed) variables on the
left, and _independent_ variables (predictors) on the right.  This formula is
then parsed by R and used to transform your data into a **model matrix**, which
can be used to fit a regression model (etc.).

```{r}

d %>%
  select(featureCorrelation, taskTransition, congruency) %>%
  head() %>%
  print()

model.matrix(log_RT ~ featureCorrelation + taskTransition + congruency, d) %>%
  head() %>%
  print()

```

Note that there's one column per predictor, where categorical variables have
been converted into numbers based on the contrasts we specified last week.
There's also an `(Intercept)` column which is all ones, which allows us to model
the constant, baseline `log_RT`.  The interpretation of the intercept
_coefficient_ will depend on the contrasts we use for our categorical
variables.  (Because our variables are all _sum_ coded, the intercept
coefficient corresponds to the grand mean).

# Fitting a simple linear regression

The vast majority of modeling functions in R take a formula and a `data.frame`
as input.  To fit a linear model:

```{r}

lm_fit <- lm(log_RT ~ featureCorrelation + taskTransition + congruency, d)

summary(lm_fit)

```

# Fitting a mixed effects regression

```{r}

library(lme4)

lmer_fit <- lmer(log_RT ~ featureCorrelation + taskTransition + congruency + (1|ID),
                 d)

summary(lmer_fit)

```
